



\vzmstitle[
	\footnote{Research was partially
supported by the grant from Russian Science Foundation (RSF) N20-11-20085}
]{
	Convergence rate when approximating fractional equations
}
\vzmsauthor{Piskarev}{S.}
\vzmsinfo{Moscow, MSU; {\it piskarev@gmail.com}}
%\vzmsauthor{Петров}{П.\,П.}
%\vzmsinfo{Урюпинск, УрГУ; {\it tractorist\_petr@example.org}}

\vzmscaption



This talk is devoted to  the  well-posedness and approximation for nonhomogeneous
fractional differential equations in Banach spaces $E:$
$$ D^\alpha u(t) = A u(t) + f(t), u(0)= u^0, $$
where $0 < \alpha <1$ and operator $A$ generates analytic $\alpha$-resolution family.
 First of all, the same way
as for integer derivative, we get the necessary and sufficient condition for the
well-posedness of   nonhomogeneous fractional Cauchy problems in the Holder spaces
$C^\beta_0([0,T];E).$ Secondly, by using implicit difference scheme and
explicit  difference scheme, we deal with the full discretization of the solutions
of nonhomogeneous fractional differential equations in time variables, get the  stability
of the implicit difference scheme and
explicit difference scheme. Finally, the order of convergence when approximating
homogeneous fractional differential equations by implicit difference scheme and
explicit difference scheme in the space $C([0,T];E)$ were obtained.




% Оформление списка литературы
% \litlist


1. {\it Li Liu, Fan Zhenbin, Li Gang, Piskarev S.} Discrete almost maximal regularity and stability for 
fractional differential equations in $L^p ([0, 1], \Omega).$ //  Applied Mathematics and Computation. - 2021, - Т. 386. 

2. {\it Ru Liu, Piskarev Sergey.} Well-Posedness and Approximation for Nonhomogeneous Fractional Differential Equations
// Numerical Functional Analysis and Optimization, издательство Marcel Dekker Inc. - 2021, - Т. 42. - № 6, - С. 619-643.



