
\begin{center}
    {\bf APPROXIMATION OF FRACTIONAL DIFFERENTIAL EQUATIONS IN BANACH SPACES\footnote{This work was supported by Russian Foundation for Basic Research
(Grant Nos. 20-51-50002 and  20-01-00015).}}\\

    {\it S. Piskarev}

    (Москва; {\it piskarev@gmail.com})
\end{center}

\addcontentsline{toc}{section}{Piskarev S.}

In this talk we have a deal with the  well-posedness and approximation for nonhomogeneous
 fractional differential equations in Banach spaces $E$:
 $$ (\mathbf{D}_{t}^{\alpha} u)(t) = A u(t) + f (t), \quad t \in [0,T];\quad
u(0)=x, $$
where $\mathbf{D}_{t}^{\alpha}$ is the Caputo-Dzhrbashyan derivative $0 < \alpha < 1.$
Follow [1] we get the necessary and sufficient condition for the well-posedness of
nonhomogeneous fractional Cauchy problems in the spaces $C_0^\beta([0,T];E)$.
Then by using implicit difference scheme and explicit difference scheme, we
deal with the full discretization of the solutions of nonhomogeneous
fractional differential equations in time variables and the same way as in [2] we
get the  stability of the schemes and the order of convergence.

We discuss also discretization of semilinear problem.


% Оформление списка литературы
\smallskip \centerline {\bf Bibliography} \nopagebreak

1. {\it Krein S.} Linear differential equations in Banach space. М.: Nauka, 1987. 408 с.

2. {\it R. Liu, Miao Li, S. Piskarev.} The Order of Convergence of Difference Schemes for
Fractional Equations. {\em Numerical Functional Analysis and Optimization}, {\bf 38} (6)
(2017), 754---769.



